Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-04-21 thesis_chapter_3.Rmd
version 3 of draft, vector map script updated

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors, and are regarded as the tumble-weed of the ocean. Despite many studies that exist on the topic, the relative role of wind and surface currents in macroalgae trajectory is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been exponential in recent years, and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
---

@@

```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}

@@

p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
p_RK4_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H100_W00_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H095_W05_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H090_W10_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H085_W15_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn). Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 2C-D). The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles. Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag. Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F). Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy. Similar behaviour is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag. In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles travelled initially southward and then westward offshore and also becoming entrained in an eddy. In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G). Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles. This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location. In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=8, fig.cap="Figure 3: Comparison of mean trajectories. Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April. Panel (E) represents hydrodynamic drag only simulations and Panel (F) represents simulations for both wind and hydrodynamic drag, for July. Panel (G) represents hydrodynamic drag only simulations and Panel (H) represents simulations for both wind and hydrodynamic drag, for October. In all panels the passive simulations are represented by black lines."}

ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("right"), labels = "AUTO", align = "hv")

@@

heatmap_Mean_KFloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("17", "18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("30","31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

leg_Oct <- get_legend(heatmap_RK4_Oct)

ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("43","44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05). The simulations that considered for 100% hydrodynamic drag travelled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations. When comparing distances between passive and kelp particles for April, significant differences are evident. (Figure 11, simulations 17-29; p < 0.05). Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled shorter distance across compared to other kelp particles. The shortest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios. No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05). In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05). Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled shorter distance compared to all other particles. As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the shortest distance compared to all other particles.

@@

ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)

@@

The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix). This was also reflected when comparing distance travelled by particles and particle density plots (see appendix). The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).

@@

Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, no research exists to clarify the drifting characteristics of solitary macroalgae. Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes. The findings from this study indicate that solitary floating virtual *E. maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle. Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies. Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E. maxima*. In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy. Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not. This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately. Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally. The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.

@@

# Make mean trajectory plot
shape_plot_April <- ggplot()+
  #geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = sphere_mean_April, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") +
  ggtitle("April") +
  #guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

>>>>

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape South African Observation Network]
output:
  word_document:
    reference_docx: markdown/template.docx
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors, and are regarded as the tumble-weed of the ocean. Despite many studies that exist on the topic, the relative role of wind and surface currents in macroalgae trajectory is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been exponential in recent years, and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
---

@@

```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}

@@

p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
p_RK4_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H100_W00_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H095_W05_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H090_W10_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H085_W15_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_Kfloat_H100_W00_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H095_W05_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H090_W10_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H085_W15_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H100_W00_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H095_W05_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H090_W10_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H085_W15_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn). Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 2C-D). The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles. Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag. Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F). Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy. Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag. In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy. In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G). Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles. This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location. In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=7.5, fig.cap="Figure 3: Comparison of mean trajectories. Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April. Panel (E) represents hydrodynamic drag only simulations and Panel (F) represents simulations for both wind and hydrodynamic drag, for July. Panel (G) represents hydrodynamic drag only simulations and Panel (H) represents simulations for both wind and hydrodynamic drag, for October. In all panels the passive simulations are represented by black lines."}

ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")

@@

heatmap_Mean_KFloat_H090_W100 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("14", "15","16","17","18","19","20","21","22","23","24","25", "26"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("27","28","29","30","31","32","33","34","35","36","37","38","39"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

leg_Oct <- get_legend(heatmap_Min_Kfloat_H100_W00_Oct)

ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Mean_KFloat_H090_W100, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("40","41","42","43","44","45","46","47","48","49","50","51","52"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))

@@

There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; p < 0.05). The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations. When comparing distances between passive and kelp particles for April, significant differences are evident. (Figure 11, simulations 17-29; p < 0.05). Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles. The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios. No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05). In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05). Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles. As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure traveled the shortest distance compared to all other particles.

@@

ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "auto", ncol = 1, nrow = 4)

@@

The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix). This was also reflected when comparing distance traveled by particles and particle density plots (see appendix). The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).

@@

Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, no research exists to clarify the drifting characteristics of solitary macroalgae. Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes. The findings from this study indicate that solitary floating virtual *E. maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle. Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies. Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E. maxima*. In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy. Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not. This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately. In addition, the results from this study highlights the importance of regional oceanographic process and how these vary seasonally to better understand how kelp flotsam is dispersed in the ocean. In particular, the current study highlights the role of the Benguela Jet and mesoscale processes and how these vary seasonally; ultimately determining transport success around the coast. Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally. The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.

@@

# Make mean trajectory plot
shape_plot_April <- ggplot()+
  #geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = sphere_mean_April, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'red')+
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") +
  ggtitle("April") +
  #guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r vector maps, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: The average ocean (left column) and wind vector maps (right column) for the respective seasons are presented. The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring). A represents summer, B autumn, C winter and D spring", fig.height=10, fig.width=8}
